Confidence Intervals for Prevalence Estimates from Complex Surveys with Imperfect Assays

### 2021-12-07 PrevAdjSurvey.tex

Each estimated component in Equation~\ref{eq:srs-beta-est} is a binomial probability parameter.
For each of these, we use distributions associated with  the exact binomial confidence interval.
For a binomial experiment with \( x \) successes out of \( n \) trials, the lower confidence distribution \( (B^L) \) is \( \operatorname{Beta}(x, n - x + 1) \), and the upper confidence distribution \( (B^U) \) is \( \operatorname{Beta}(x + 1, n - x)\).
Let \( q(a, W) \) be the \( a \)th quantile of a random variable \( W \). Then the exact \( 1 - \alpha \)\% central confidence interval of Clopper-Pearson for the binomial parameter is 
\begin{equation}
\left\{ q \left( \frac{\alpha}{2}, B^L \right), q \left( 1 - \frac{\alpha}{2}, B^U, \right) \right\}.
\label{eq:C-P}
\end{equation}

@@

where analogously to the Clopper-Pearson interval (see equation~\ref{eq:C-P}),
\begin{equation*}
    % B^L_{KG} \sim \operatorname{Beta}\left(n_{\text{eff}} \sum_{i=1}^k w_i \hat{\theta}_i, n_{\text{eff}} \left(1 -  \sum_{i=1}^k w_i \hat{\theta}_i\right) + 1\right)
    B^L_{KG} \sim \operatorname{Beta}\left(x_{\text{eff}},  n_{\text{eff}} -  x_{\text{eff}} + 1 \right)
\end{equation*}

@@

We compare the wsPoisson method to the more traditional Agresti-Coull and Korn-Graubard methods for survey proportions in a variety of settings.
Our simulations examine varying levels of disease prevalence (0.5\% or 5\%), types of groups surveyed (50 groups 200 subjects or 8000 groups of 1 subject), distributions of weights among the groups (coefficient of variation from approximately 0\% to nearly 6\%), the number of groups with non-zero prevalence, and the weights of the groups with non-zero prevalence.
For each scenario, 10,000 data sets are simulated and assessed.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_coverage_50_groups_0_005_prev.pdf}
\caption{Coverage properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal coverage, 95\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_coverage_50_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_coverage_8000_groups_0_005_prev.pdf}
\caption{Coverage properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal coverage, 95\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_coverage_8000_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_coverage_50_groups_0_05_prev.pdf}
\caption{Coverage properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal coverage, 95\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_coverage_50_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_coverage_8000_groups_0_05_prev.pdf}
\caption{Coverage properties for the wsPoisson model and two standard methods, Agresti-Coull and Korn-Graubard.
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal coverage, 95\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_coverage_8000_groups_0_05_prev}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_lower_error_frequency_50_groups_0_005_prev.pdf}
\caption{Lower error properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal lower error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_lower_error_frequency_50_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_lower_error_frequency_8000_groups_0_005_prev.pdf}
\caption{Lower error properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal lower error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_lower_error_frequency_8000_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_lower_error_frequency_50_groups_0_05_prev.pdf}
\caption{Lower error properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal lower error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_lower_error_frequency_50_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_lower_error_frequency_8000_groups_0_05_prev.pdf}
\caption{Lower error properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal lower error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_lower_error_frequency_8000_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_upper_error_frequency_50_groups_0_005_prev.pdf}
\caption{Upper error properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal upper error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_upper_error_frequency_50_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_upper_error_frequency_8000_groups_0_005_prev.pdf}
\caption{Upper error properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal upper error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_upper_error_frequency_8000_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_upper_error_frequency_50_groups_0_05_prev.pdf}
\caption{Upper error properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal upper error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_upper_error_frequency_50_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_upper_error_frequency_8000_groups_0_05_prev.pdf}
\caption{Upper error properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal upper error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_upper_error_frequency_8000_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_confidence_interval_width_50_groups_0_005_prev}
\caption{Confidence interval width properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 50 groups of 200 people are sampled.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_confidence_interval_width_50_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_confidence_interval_width_50_groups_0_05_prev}
\caption{Confidence interval width properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 50 groups of 200 people are sampled.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_confidence_interval_width_50_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_confidence_interval_width_8000_groups_0_005_prev}
\caption{Confidence interval width properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 8000 individuals are sampled.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_confidence_interval_width_8000_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_confidence_interval_width_8000_groups_0_05_prev}
\caption{Confidence interval width properties for the wsPoisson model and two standard methods, Agresti-Coull and Clopper-Pearson.
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 8000 individuals are sampled.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_confidence_interval_width_8000_groups_0_05_prev}
\end{figure}

>>>>

Each estimated component in Equation~\ref{eq:srs-beta-est} is a binomial probability parameter.
For each of these, we use distributions associated with  the exact binomial confidence interval.
For a binomial experiment with \( x \) successes out of \( n \) trials, the lower confidence distribution \( (B^L) \) is \( \operatorname{Beta}(x, n - x + 1) \), and the upper confidence distribution \( (B^U) \) is \( \operatorname{Beta}(x + 1, n - x)\).
Let \( q(a, W) \) be the \( a \)th quantile of a random variable \( W \). Then the exact \( 1 - \alpha \)\% central confidence interval of Korn-Graubard for the binomial parameter is 
\begin{equation}
\left\{ q \left( \frac{\alpha}{2}, B^L \right), q \left( 1 - \frac{\alpha}{2}, B^U, \right) \right\}.
\label{eq:C-P}
\end{equation}

@@

where analogously to the Korn-Graubard interval (see equation~\ref{eq:C-P}),
\begin{equation*}
    % B^L_{KG} \sim \operatorname{Beta}\left(n_{\text{eff}} \sum_{i=1}^k w_i \hat{\theta}_i, n_{\text{eff}} \left(1 -  \sum_{i=1}^k w_i \hat{\theta}_i\right) + 1\right)
    B^L_{KG} \sim \operatorname{Beta}\left(x_{\text{eff}},  n_{\text{eff}} -  x_{\text{eff}} + 1 \right)
\end{equation*}

@@

We compare the wsPoisson method to the more traditional Dean-Pagano modification of the Agresti-Coull (DPAC) method and the Korn-Graubard (KG) method for survey proportions in a variety of settings.
Our simulations examine varying levels of disease prevalence (0.5\% or 5\%), types of groups surveyed (50 groups 200 subjects or 8000 groups of 1 subject), distributions of weights among the groups (coefficient of variation from approximately 0\% to nearly 6\%), the number of groups with non-zero prevalence, and the weights of the groups with non-zero prevalence.
For each scenario, 10,000 data sets are simulated and assessed.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_coverage_50_groups_0_005_prev.pdf}
\caption{Coverage properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal coverage, 95\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_coverage_50_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_coverage_8000_groups_0_005_prev.pdf}
\caption{Coverage properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal coverage, 95\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_coverage_8000_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_coverage_50_groups_0_05_prev.pdf}
\caption{Coverage properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal coverage, 95\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_coverage_50_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_coverage_8000_groups_0_05_prev.pdf}
\caption{Coverage properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal coverage, 95\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_coverage_8000_groups_0_05_prev}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_lower_error_frequency_50_groups_0_005_prev.pdf}
\caption{Lower error properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal lower error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_lower_error_frequency_50_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_lower_error_frequency_8000_groups_0_005_prev.pdf}
\caption{Lower error properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal lower error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_lower_error_frequency_8000_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_lower_error_frequency_50_groups_0_05_prev.pdf}
\caption{Lower error properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal lower error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_lower_error_frequency_50_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_lower_error_frequency_8000_groups_0_05_prev.pdf}
\caption{Lower error properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal lower error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_lower_error_frequency_8000_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_upper_error_frequency_50_groups_0_005_prev.pdf}
\caption{Upper error properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal upper error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_upper_error_frequency_50_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_upper_error_frequency_8000_groups_0_005_prev.pdf}
\caption{Upper error properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal upper error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_upper_error_frequency_8000_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_upper_error_frequency_50_groups_0_05_prev.pdf}
\caption{Upper error properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 50 groups of 200 people are sampled.
The horizontal dashed line indicates the nominal upper error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_upper_error_frequency_50_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_upper_error_frequency_8000_groups_0_05_prev.pdf}
\caption{Upper error properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 8000 individuals are sampled.
The horizontal dashed line indicates the nominal upper error rate, 2.5\%.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_upper_error_frequency_8000_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_confidence_interval_width_50_groups_0_005_prev}
\caption{Confidence interval width properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 50 groups of 200 people are sampled.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_confidence_interval_width_50_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_confidence_interval_width_50_groups_0_05_prev}
\caption{Confidence interval width properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 50 groups of 200 people are sampled.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_confidence_interval_width_50_groups_0_05_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_confidence_interval_width_8000_groups_0_005_prev}
\caption{Confidence interval width properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 0.5\% Prevalence where 8000 individuals are sampled.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_confidence_interval_width_8000_groups_0_005_prev}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/perfect_confidence_interval_width_8000_groups_0_05_prev}
\caption{Confidence interval width properties for the wsPoisson model and two standard methods, the Dean-Pagano modification of Agresti-Coull and Korn-Graubard
Each point represents 10,000 simulations of datasets from a population with 5\% Prevalence where 8000 individuals are sampled.
Colored dashed lines are estimates from a logistic regression model using cubic splines.}
\label{fig:perfect_confidence_interval_width_8000_groups_0_05_prev}
\end{figure}
